---
title: "Your PRISMA flowchart"
output: word_document
geometry: margin = 2mm
always_allow_html: yes
---

```{r setup, include=FALSE}
library("tidyverse", "readxl", "here")

knitr::opts_chunk$set(echo = TRUE)
version <- 1

```

This script generates a PRISMA 2020 flow Diagram from Page MJ, McKenzie JE, Bossuyt PM, Boutron I, Hoffmann TC, Mulrow CD, et al. [The PRISMA 2020 statement: an updated guideline for reporting systematic reviews](http://doi.org/10.1136/bmj.n71). BMJ 2021;372:n71. doi: 10.1136/bmj.n71

The script is inspired by Peter Higgins's tutorial from 26th January 2019, available here: https://rpubs.com/phiggins/461686

```{r read_data, message=FALSE, warning=FALSE, include=FALSE}

PRISMA_data <- read.csv(here::here("PRISMA.csv"))
```

```{r wrangle_data, message=FALSE, warning=FALSE, include=FALSE}

PRISMA_data$recScreened <- PRISMA_data$records_db + PRISMA_data$handSearch - PRISMA_data$duplRec - PRISMA_data$autoIneligble - PRISMA_data$remOther
PRISMA_data$recRetr <- PRISMA_data$recScreened - PRISMA_data$screenRecExcl
PRISMA_data$repAssessed <- PRISMA_data$recRetr - PRISMA_data$recNotRetr
PRISMA_data$reportIncl <- PRISMA_data$repAssessed - PRISMA_data$exclReason1 - PRISMA_data$exclReason2 - PRISMA_data$exclReason3 - PRISMA_data$exclReason4 - PRISMA_data$exclReason6 - PRISMA_data$exclReason6 - PRISMA_data$exclReason7 

```

```{r flowchart_canvas, message=FALSE, warning=FALSE, include=FALSE}
#Creating a working space. Grid lines are to help to align and place boxes.

#Define the size of canvas
tibble_min = 1
tibble_max = 150 #increase this if the lower part of the chart gets cut-off

#CONSORT
data <- tibble(x= tibble_min:tibble_max, y= tibble_min:tibble_max)

data %>% 
  ggplot(aes(x, y)) +
  scale_x_continuous(minor_breaks = seq(10, tibble_max, 10)) +
  scale_y_continuous(minor_breaks = seq(10, tibble_max, 10)) +
  labs(caption = "Created using RMarkdown script by Vainre (2021)") +
  theme_void() -> #for easier editing, replace this with theme_linedraw()
  p


```

```{r flowchart_parameters, message=FALSE, warning=FALSE, include=FALSE}
#setting quality and dimensions for the image. Reduce width and height if you need more space or need to be more conservative with space
knitr::opts_chunk$set(dpi=800, fig.width = 11, fig.height = 9)

#the parameters for the flowchart
gap_bw_columns <- 10

x_col1_start <- tibble_min
x_col1_end <- tibble_max/2-gap_bw_columns
x_col1_mid <- (x_col1_end-x_col1_start)/2

x_col2_start <- x_col1_end+2*gap_bw_columns
x_col2_end <- tibble_max
x_col2_mid <- (x_col2_end-x_col2_start)/2+x_col2_start

#what's the height of a box that contains just 1 line of text?
height_y_1line <- 6 

# specify where rows start
row1 <- tibble_max                #records identified
row2 <- row1 - 8*height_y_1line   #records screened
row3 <- row2 - 4*height_y_1line   #records sought for retrieval
row4 <- row3 - 4*height_y_1line   #reports assessed for eligibility
row5 <- row4 - 4*height_y_1line   #records included in review

#Aesthetics
bordercolour <- "grey" #the colour of the borders
arrowcolour <- bordercolour #colour of the arrows (line + end)
fillcolour <- "white" #the colour of boxes
textsize <- 5.5
border_width <- 0.75
arrow_width <- border_width
arrow_size <- 3
line_spacing <- 1 

#rm(gap_bw_columns, tibble_min, tibble_max)
```

```{r generate_flowchart, echo=FALSE, message=FALSE, warning=FALSE}
########################  IDENTIFICATION ######################## 

# Texts searched through databases
p +
  geom_rect(xmin = x_col1_start, 
            xmax = x_col1_end, 
            ymin = row1-3*height_y_1line, 
            ymax = row1, 
            colour = bordercolour,
            fill = fillcolour, 
            size = border_width ) +
  annotate('text', 
           x = x_col1_mid, 
           y = row1-3*height_y_1line/2,
           label = paste("Records identified from\n Databases: k =", PRISMA_data$records_db,"\n Hand searches: k =", PRISMA_data$handSearch),            
           size = textsize) ->
  p

#Records removed before screening

p +
  geom_rect(xmin = x_col2_start, 
            xmax = x_col2_end, 
            ymin = row1-6*height_y_1line, 
            ymax = row1, 
            colour = bordercolour,
            fill = fillcolour, 
            size = border_width ) +
  annotate('text', 
           x = x_col2_mid, 
           y = row1-3*height_y_1line,
           label = paste("Records removed before screening\n Duplicate records: k =", PRISMA_data$duplRec,"\n Records marked as ineligible\n by automation tools: k =", PRISMA_data$autoIneligble, "\n Records removed for other reasons:\n k =", PRISMA_data$remOther),            
           size = textsize) ->
  p

#Row1 -> row 2 arrows
#horisontal
p +
  geom_segment(x = x_col1_end, 
               xend = x_col2_start, 
               y = row1-1.5*height_y_1line,
               yend = row1-1.5*height_y_1line, 
               size = arrow_width,
               colour = arrowcolour,
               linejoin = "mitre", 
               lineend = "butt",
               arrow = arrow(length = unit(arrow_size, "mm"),
                             type= "closed")) +
  #down
  geom_segment(x = x_col1_mid, 
               xend = x_col1_mid,
               y = row1-3*height_y_1line,
               yend = row2,
               size = arrow_width, 
               colour = arrowcolour, 
               linejoin = "mitre", 
               lineend = "butt",
               arrow = arrow(length = unit(arrow_size, "mm"),
                             type= "closed")) ->
  p


######################## SCREENING ######################## 

#Records screened
p +
  geom_rect(xmin = x_col1_start, 
            xmax = x_col1_end, 
            ymin = row2-2*height_y_1line, 
            ymax = row2, 
            colour = bordercolour,
            fill = fillcolour, 
            size = border_width ) +
  annotate('text', 
           x = x_col1_mid, 
           y = row2-height_y_1line,
           label = paste("Records screened:\n k =", PRISMA_data$recScreened), 
           size = textsize) ->
  p


#Records excluded
p +
  geom_rect(xmin = x_col2_start, 
            xmax = x_col2_end, 
            ymin = row2-2*height_y_1line, 
            ymax = row2, 
            colour = bordercolour,
            fill = fillcolour, 
            size = border_width ) +
  annotate('text', 
           x = x_col2_mid, 
           y = row2-2*height_y_1line/2,
           label = paste("Records excluded:\n k =", PRISMA_data$screenRecExcl), 
           size = textsize) ->
  p


#Row2 -> row 3 arrows
#horisontal
p +
  geom_segment(x = x_col1_end, 
               xend = x_col2_start, 
               y = row2-height_y_1line,
               yend = row2-height_y_1line, 
               size = arrow_width, 
               colour = arrowcolour, 
               linejoin = "mitre", 
               lineend = "butt",
               arrow = arrow(length = unit(arrow_size, "mm"),
                             type= "closed")) +
  #down
  geom_segment(x = x_col1_mid, 
               xend = x_col1_mid,
               y = row2 - 2*height_y_1line, #make sure your height_y is correct
               yend = row3,
               size = arrow_width, 
               colour = arrowcolour, 
               linejoin = "mitre", 
               lineend = "butt",
               arrow = arrow(length = unit(arrow_size, "mm"),
                             type= "closed")) ->
  p

#Reports sought for retrieval
p +
  geom_rect(xmin = x_col1_start, 
            xmax = x_col1_end, 
            ymin = row3-2*height_y_1line, 
            ymax = row3, 
            colour = bordercolour,
            fill = fillcolour, 
            size = border_width ) +
  annotate('text', 
           x = x_col1_mid, 
           y = row3-height_y_1line,
           label = paste("Reports sought for retrieval:\n k =", PRISMA_data$recRetr), 
           size = textsize) ->
  p

#Records not retreived
p +
  geom_rect(xmin = x_col2_start, 
            xmax = x_col2_end, 
            ymin = row3-2*height_y_1line, 
            ymax = row3, 
            colour = bordercolour,
            fill = fillcolour, 
            size = border_width ) +
  annotate('text', 
           x = x_col2_mid, 
           y = row3-height_y_1line,
           label = paste("Reports not retreived:\n k =", PRISMA_data$recNotRetr), 
           size = textsize) ->
  p

#Row3 -> row 4 arrows
#horisontal
p +
  geom_segment(x = x_col1_end, 
               xend = x_col2_start, 
               y = row3-height_y_1line,
               yend = row3-height_y_1line, 
               size = arrow_width, 
               colour = arrowcolour, 
               linejoin = "mitre", 
               lineend = "butt",
               arrow = arrow(length = unit(arrow_size, "mm"),
                             type= "closed")) +
  #down
  geom_segment(x = x_col1_mid, 
               xend = x_col1_mid,
               y = row3-2*height_y_1line, #make sure your height_y is correct, should match the height of the box
               yend = row4,
               size = arrow_width, 
               colour = arrowcolour, 
               linejoin = "mitre", 
               lineend = "butt",
               arrow = arrow(length = unit(arrow_size, "mm"),
                             type= "closed")) ->
  p


#Reports assessed for eligibility
p +
  geom_rect(xmin = x_col1_start, 
            xmax = x_col1_end, 
            ymin = row4-2*height_y_1line, 
            ymax = row4, 
            colour = bordercolour,
            fill = fillcolour, 
            size = border_width ) +
  annotate('text', 
           x = x_col1_mid, 
           y = row4-height_y_1line,
           label = paste("Reports assessed for eligibility:\n k =", PRISMA_data$repAssessed), 
           size = textsize) ->
  p

#Reports excluded
p +
  geom_rect(xmin = x_col2_start, 
            xmax = x_col2_end, 
            ymin = row4-6*height_y_1line, 
            ymax = row4, 
            colour = bordercolour,
            fill = fillcolour, 
            size = border_width ) +
  annotate('text', 
           x = x_col2_mid, 
           y = row4-3*height_y_1line,
           label = paste("Reports excluded:\n Reason 1 =", PRISMA_data$exclReason1,"\n Reason 2 =", PRISMA_data$exclReason2,"\n Reason 3 =", PRISMA_data$exclReason3,"\n Reason 4 =", PRISMA_data$exclReason4,"\n Reason 5 =", PRISMA_data$exclReason5), 
           size = textsize) ->
  p


#Row4 -> row 5 arrows
#horisontal
p +
  geom_segment(x = x_col1_end, 
               xend = x_col2_start, 
               y = row4-height_y_1line,
               yend = row4-height_y_1line, 
               size = arrow_width, 
               colour = arrowcolour, 
               linejoin = "mitre", 
               lineend = "butt",
               arrow = arrow(length = unit(arrow_size, "mm"),
                             type= "closed")) +
  #down
  geom_segment(x = x_col1_mid, 
               xend = x_col1_mid,
               y = row4 - 2*height_y_1line, #make sure your height_y is correct, should match the height of the box
               yend = row5,
               size = arrow_width, 
               colour = arrowcolour, 
               linejoin = "mitre", 
               lineend = "butt",
               arrow = arrow(length = unit(arrow_size, "mm"),
                             type= "closed")) ->
  p

########################  INCLUDED  ######################## 

#Reports included in review
p +
  geom_rect(xmin = x_col1_start, 
            xmax = x_col1_end, 
            ymin = row5-4*height_y_1line, 
            ymax = row5, 
            colour = bordercolour,
            fill = fillcolour, 
            size = border_width ) +
  annotate('text', 
           x = x_col1_mid, 
           y = row5-2*height_y_1line,
           label = paste("Studies included in review:\n k =", PRISMA_data$studyIncl,"\n Reports included in review:\n k =", PRISMA_data$reportIncl), 
           size = textsize) ->
  p

ggsave(plot = p, "PRISMA2020.png")

p
```

Please cite as: 
Vainre, Maris. (2021).  RMarkdown PRISMA flowchart generator for Word. Version `r version`. GitHub: doi
